---
title: "POA Accidents Plots"
output: pdf_document
---



#Download

```{r}
library(tidyverse)
file = "acidentes-2016.csv"
if(!file.exists(file)){
  download.file("http://datapoa.com.br/storage/f/2017-08-03T13%3A19%3A45.538Z/acidentes-2016.csv",
	destfile=file)
}
```


#Is there a time of the year with more accidents?



How do I represent a time (or interval) of year. A simple solution would be by month, but I decided to be more detailed and did by date. Date is represented in the DATA column. So, first I convert the DATA column to the date format, then I count the number of accidents per date.

```{r}
df %>%
  select(DATA) %>%
  mutate(DATA = as.Date(DATA, format = "%Y-%m-%d")) %>%
  group_by(DATA) %>%
  summarize(NumberOfAccidents = n())   -> df.accidentsByDate;
df.accidentsByDate;
```

If I plot this number, there will be too much noise in the plot. Instead, I smooth the data a little. I should use a gaussian filter, but since I don't know how yet, I will simply perform a rolling average of each 7 day period.

```{r}
df.accidentsByDate %>%
   mutate(smooth = (NumberOfAccidents
          + lag(NumberOfAccidents,1) 
          + lag(NumberOfAccidents,2)
          + lag(NumberOfAccidents,3)
          + lead(NumberOfAccidents,1)
          + lead(NumberOfAccidents,2)
          + lead(NumberOfAccidents,3)
          )/7) ->df.accidentsByDate


df.accidentsByDate %>% 
  ggplot(aes(x=DATA,y=smooth)) + 
  geom_line() + 
  scale_x_date("Days",date_breaks = "1 month", date_labels = "%b") +
  ylab("7-day rolling average")
```

This has problems, since it discards the edges of the dataframe, but its adequate.


#How many vehicles are usually involved?

There are many types of vehicles that are separated by collumns in the dataset. I will try to create a new collumn that aggregates the total number of vehicles in an accident. Then I simply plot the number of instances of each total number of vehicles in a histogram.

```{r}
df %>%
mutate(TOTALVEHICLES = AUTO + TAXI + LOTACAO + ONIBUS_URB + ONIBUS_MET + ONIBUS_INT + CAMINHAO + MOTO + CARROCA + BICICLETA + OUTRO) -> df.totalVehicles; 


df.totalVehicles %>% 
  select(TOTALVEHICLES) %>%
  ggplot(aes(x=TOTALVEHICLES)) + 
  geom_bar() + 
  geom_text(stat='count',  aes(label=..count..), nudge_y = 250) +
  scale_x_continuous("Number of Vehicles", breaks = c(0,1,2,3,4,5,6,7) ) + 
  ylab("Number of Accidents") 

```

Curiously, there is an accident with 0 vehicles.

#Is there a specific weekday with more accidents?

The weekday is defined in the dataset as a factor. Counting the number for each weekday is simple, but the order of the factors is wrong. I first estabilish this order, then I plot the number of instances of each weekday. I also swap the coordinates, so that the weekdays can be read more easily. 

```{r}
levels(df$DIA_SEM) <- c("DOMINGO", "SEGUNDA-FEIRA", "TERCA-FEIRA", "QUARTA-FEIRA", "QUINTA-FEIRA", "SEXTA-FEIRA", "SABADO")

df.accidentsByWeekDay <- df %>%
  select(DIA_SEM)

df.accidentsByWeekDay %>% 
  ggplot(aes(x=DIA_SEM, fill=DIA_SEM)) + 
  geom_bar() + 
  geom_text(stat='count',  aes(label=..count..), nudge_y = -100) +
  scale_fill_brewer(palette = "Dark2", guide=FALSE) +
  theme_light() +
  scale_y_continuous(limits = c(0,2200), expand = c(0,0)) +
  xlab("Weekday") + 
  ylab("Total Number of Accidents") + 
  coord_flip()
  
```


