---
title: "POA Accidents"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#Download

```{r}
library(tidyverse)
file = "acidentes-2016.csv"
if(!file.exists(file)){
  download.file("http://datapoa.com.br/storage/f/2017-08-03T13%3A19%3A45.538Z/acidentes-2016.csv",
	destfile=file)
}
```

#Preliminary data analysis

```{r}
df <- read.csv(file, header=TRUE, sep=";");
df;
```


```{r}
 df %>%  str
```

```{r}
 df %>%  summary
```

#Is there a time of the year with more accidents?

I will try to analyse the number of accidents by month. First I will select the data:

```{r}
Months <-  as.factor(c("JAN", "FEV", "MAR", "APR", "MAY", "JUN", "JUL", "AGO", "SEP", "OCT", "NOV", "DEC"))

df %>%
  select(MES) %>%
  group_by(MES) %>%
  summarize(NumberOfAccidents = n()) %>%
  mutate(MonthAbr = Months[MES]) %>%
  select(MonthAbr, NumberOfAccidents)-> df.accidentsByMonth;
df.accidentsByMonth;
levels(df.accidentsByMonth$MonthAbr) <-c("JAN", "FEV", "MAR", "APR", "MAY", "JUN", "JUL", "AGO", "SEP", "OCT", "NOV", "DEC") ;
levels(df.accidentsByMonth$MonthAbr) ;
```


I also want to try to plot something:

```{r}
  df.accidentsByMonth %>% ggplot(aes(x=MonthAbr, y = NumberOfAccidents, fill = MonthAbr )) + 
  geom_col() + 
  #geom_text(aes(label=..count..)) +
  #scale_fill_brewer(palette = "SAT1", guide=FALSE)
  ylab("Total Number of Accidents")
```


How about by date? First I convert the DATA column to the date format, then I count the number of accidents per date.

```{r}
df %>%
  select(DATA) %>%
  mutate(DATA = as.Date(DATA, format = "%Y-%m-%d")) %>%
  group_by(DATA) %>%
  summarize(NumberOfAccidents = n())   -> df.accidentsByDate;
df.accidentsByDate;
```

Then I plot the previous dataframe: 

```{r}
  df.accidentsByDate %>% 
  ggplot(aes(x=DATA,y=NumberOfAccidents)) + 
  geom_line()
```

There's too much noise. Let's smooth the data a little. I should use a gaussian filter, but since I don't know how yet, I will simply average each 7 day period.

```{r}
df.accidentsByDate %>%
   mutate(smooth = (NumberOfAccidents
          + lag(NumberOfAccidents,1) 
          + lag(NumberOfAccidents,2)
          + lag(NumberOfAccidents,3)
          + lead(NumberOfAccidents,1)
          + lead(NumberOfAccidents,2)
          + lead(NumberOfAccidents,3)
          )/7) ->df.accidentsByDate
df.accidentsByDate;

df.accidentsByDate %>% 
  ggplot(aes(x=DATA,y=smooth)) + 
  geom_line() + 
  scale_x_date("Days",date_breaks = "1 month", date_labels = "%b") +
  ylab("7-day rolling average")
```
This has problems, since it discards the edges of the dataframe, but its adequate.


#How many vehicles are usually involved?

There are many types of vehicles that are separated by collumns in the dataset. I will try to create a new collumn that aggregates the total number of vehicles in an accident.

```{r}
df %>%
mutate(TOTALVEHICLES = AUTO + TAXI + LOTACAO + ONIBUS_URB + ONIBUS_MET + ONIBUS_INT + CAMINHAO + MOTO + CARROCA + BICICLETA + OUTRO) -> df.totalVehicles; 
df.totalVehicles %>% select(TOTALVEHICLES);
df.totalVehicles %>% select(TOTALVEHICLES) %>% summary()
```

```{r}
df.totalVehicles %>% 
  select(TOTALVEHICLES) %>%
  ggplot(aes(x=TOTALVEHICLES)) + 
  geom_bar() + 
  geom_text(stat='count',  aes(label=..count..), nudge_y = 250) +
  #scale_fill_brewer(palette = "Dark2", guide=FALSE) +
  scale_x_continuous("Number of Vehicles", breaks = c(0,1,2,3,4,5,6,7) ) + 
  ylab("Number of Accidents") 
#hist(df1$TOTALVEHICLES, breaks=10, xlab="Z", main="Total Vehicles")
```

There are accidents with 0 vehicles?
```{r}
df.totalVehicles %>% filter(TOTALVEHICLES ==0)
```


Weird.

#Is there a specific weekday with more accidents?

```{r}
levels(df$DIA_SEM) <- c("DOMINGO", "SEGUNDA-FEIRA", "TERCA-FEIRA", "QUARTA-FEIRA", "QUINTA-FEIRA", "SEXTA-FEIRA", "SABADO")

df.accidentsByWeekDay <- df %>%
  select(DIA_SEM) #%>%
  #group_by(DIA_SEM) %>%
  #summarize(NumberOfAccidents = n())  
df.accidentsByWeekDay;

df.accidentsByWeekDay %>% 
  ggplot(aes(x=DIA_SEM, fill=DIA_SEM)) + 
  geom_bar() + 
  geom_text(stat='count',  aes(label=..count..)) +
  scale_fill_brewer(palette = "Dark2", guide=FALSE) +
  xlab("Weekday") + 
  ylab("Total Number of Accidents") + 
  coord_flip()
```







#Advanced: are there some area in POA with more accidents?

Simple answer: Which region has more accidents:




```{r}
df %>%
  select(REGIAO) %>%
  group_by(REGIAO) %>%
  summarize(NumberOfAccidents = n()) -> df.accidentsByRegion;
df.accidentsByRegion;
```

Another simple answer: Are there more accidents in intersections?

```{r}
df %>%
  select(LOCAL) %>%
  group_by(LOCAL) %>%
  summarize(NumberOfAccidents = n()) -> df.accidentsByLocal;
df.accidentsByLocal;
```


By full address. I rely on LOCAL_VIA being consistent:

```{r}
df %>%
  select(LOCAL_VIA) %>%
  group_by(LOCAL_VIA) %>%
  summarize(NumberOfAccidents = n()) %>% arrange(desc(NumberOfAccidents)) -> df.accidentsByFullAddress;
df.accidentsByFullAddress;
```


By road. This is more complicated, as there are two columns, with the second only being present when it is an intersection. I have to perform a join of the total number of accidents based both columns and sum that number.


```{r}
df %>%
  select(LOG1) %>%
  group_by(LOG1) %>%
  summarize(NumberOfAccidents = n()) %>%
  arrange(desc(NumberOfAccidents)) -> df.accidentsByLog1;

colnames(df.accidentsByLog1)[colnames(df.accidentsByLog1)=="LOG1"] <- "LOG"


df %>%
  select(LOG2) %>%
  group_by(LOG2) %>%
  summarize(NumberOfAccidents = n()) %>% 
  arrange(desc(NumberOfAccidents)) %>%
  filter (LOG2 != "") -> df.accidentsByLog2;

colnames(df.accidentsByLog2)[colnames(df.accidentsByLog2)=="LOG2"] <- "LOG"

df.accidentsByLog <- merge(x = df.accidentsByLog1, y = df.accidentsByLog2, by="LOG",  all = TRUE)
df.accidentsByLog[is.na(df.accidentsByLog)]<-0
df.accidentsByLog <- df.accidentsByLog %>%  mutate(NumberOfAccidents = NumberOfAccidents.x + NumberOfAccidents.y) %>% select(LOG, NumberOfAccidents) %>% arrange(desc(NumberOfAccidents))

df.accidentsByLog;

```